---
title: "Problem Sheet 2"
author: "Alexander Ober"
date: "2/22/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code is posted at the following site: https://github.com/alexanderober/BUSI525/tree/main/PS2

## Part 1

I simply follow what is suggested in the problem sheet. I've commented throughout to hopefully make it clear what I'm doing. Note the CDFs of the actual and bootstrapped match reasonably closely, which is nice. 

```{r, echo = FALSE}
## generate variables and simulated data
N <- 1000
T <- 120
mean_mkt <- 0.05/12
sd_mkt <- 0.2/sqrt(12)
sd_res <- 0.1/sqrt(12)
alpha <- rep(0, N)

market_ret <- rnorm(T, mean = mean_mkt, sd = sd_mkt)
M <- matrix(rep(market_ret, N), nrow=T)
M_res <- matrix(rnorm(N*T, mean = 0, sd = sd_res), nrow=T)
M_alpha <- matrix(rep(alpha, T), nrow = T)

## simulate returns in a T by N matrix. Each column is a fund.
M_ret <- M_alpha + M + M_res

## initialize alpha estimates, pvalues, and tstats
alpha_hat = rep(0, N)
beta_hat = rep(0, N)
pvals = rep(0, N)
tstats_orig = rep(0, N)

## matrix of redsiduals 
resid = matrix(rep(0, N*T), nrow = T)

## run a regression for each fund 
for (i in 1:N){
  Y <- M_ret[, i]
  X <- M[, i]
  
  mod <- lm(Y~X)
  summ <- summary(mod)
  
  alpha_hat[i] <- summ$coefficients[1, 1]
  beta_hat[i] <- summ$coefficients[2, 1]
  tstats_orig[i] <- summ$coefficients[1, 3]
  pvals[i] <- summ$coefficients[1, 4]
  resid[, i] <- resid(mod)
                     
}

## Bootstrap
B = 100
alpha_hat_boot = matrix(rep(0, N*B), nrow = B)
tstats_boot = matrix(rep(0, N*B), nrow = B)

for (i in 1:B){
  ## step 1
  sam <- sample(c(1:T), T, replace = TRUE)
  sor <- sort(sam)
  
  for (j in 1:N){
    
    ## step 2
    rets = beta_hat[j]*market_ret[sor] + resid[sor, j]
    
    ## step 3
    mod <- lm(rets~market_ret[sor])
    summ <- summary(mod)
    alpha_hat_boot[i, j] <- summ$coefficients[1, 1]
    tstats_boot[i, j] <- summ$coefficients[1, 3]
  }
}

## Sort for each bootstrap
for (i in 1:B){
  alpha_hat_boot[i, ] = sort(alpha_hat_boot[i, ])
  tstats_boot[i, ] = sort(tstats_boot[i, ])
}

## Compute average bootstrap distribution
distr_alpha <- colMeans(alpha_hat_boot)
distr_tstats <- colMeans(tstats_boot)

plot(ecdf(distr_tstats), 
     main = "Bootstrapped/Actual T Stat CDFs", col = 3)
lines(ecdf(tstats_orig), lty=3, col =20)
legend(x = 'topleft',legend =  c('Bootstrap','Actual'), 
       col = c(3,20), lwd = 2)

plot(ecdf(alpha_hat_boot[,50]), main = "CDF of 5% Bootstrapped
     t-stat")
plot(ecdf(alpha_hat_boot[,950]), main = "CDF of 95% Bootstrapped
     t-stat")

```

## Part 2

Note as lambda and the alpha increase, the actual CDF moves to the right relative to the bootstrapped CDF. This is what we'd expect.

```{r, echo = FALSE}

for (alpha_0 in c(0.01, 0.025, 0.05)){
  for (lambda in c(0.1, 0.25, 0.5, 0.75)){
    
    ## generate a proportion of positive alphas
    alpha <- c(rep(alpha_0/12, lambda*N), rep(0, (1-lambda)*N))
    M_alpha <- matrix(rep(alpha, T), nrow = T, byrow = TRUE)
    
    ## skilled funds correspond to the first lambda*N columns of the matrix
    M_ret <- M_alpha + M + M_res
    
    alpha_hat = rep(0, N)
    pvals = rep(0, N)
    tstats_orig = rep(0, N)
    
    ## run regressions for each fund
    for (i in 1:N){
      Y <- M_ret[, i]
      X <- M[, i]
      
      mod <- lm(Y~X)
      summ <- summary(mod)
      
      alpha_hat[i] <- summ$coefficients[1, 1]
      beta_hat[i] <- summ$coefficients[2, 1]
      tstats_orig[i] <- summ$coefficients[1, 3]
      pvals[i] <- summ$coefficients[1, 4]
      resid[, i] <- resid(mod)
      
    }
    
    ## Bootstrap
    alpha_hat_boot = matrix(rep(0, N*B), nrow = B)
    tstats_boot = matrix(rep(0, N*B), nrow = B)
    for (i in 1:B){
      
      ## step 1
      sam <- sample(c(1:T), T, replace = TRUE)
      sor <- sort(sam)
      
      for (j in 1:N){
        
        ## step 2
        rets = beta_hat[j]*market_ret[sor] + resid[sor, j]
        
        ## step 3
        mod <- lm(rets~market_ret[sor])
        summ <- summary(mod)
        alpha_hat_boot[i, j] <- summ$coefficients[1, 1]
        tstats_boot[i, j] <- summ$coefficients[1, 3]
      }
    }
    
    ## Sort for each bootstrap
    for (i in 1:B){
      alpha_hat_boot[i, ] = sort(alpha_hat_boot[i, ])
      tstats_boot[i, ] = sort(tstats_boot[i, ])
    }
    
    ## Compute average bootstrap distribution
    distr_alpha <- colMeans(alpha_hat_boot)
    distr_tstats <- colMeans(tstats_boot)
    
    ## plots
    title <- paste0("Bootstrapped/Actual T Stat CDFs, alpha_0 = ",
                    alpha_0, ", lambda = ", lambda)
    plot(ecdf(distr_tstats), main = title, col = 3)
    lines(ecdf(tstats_orig), lty=3, col =20)
    legend(x = 'topleft',legend =  c('Bootstrap','Actual'), 
           col = c(3, 20), lwd = 2)
    
    plot(ecdf(alpha_hat_boot[,50]), 
         main = "CDF of 5% Bootstrapped t-stat")
    plot(ecdf(alpha_hat_boot[,950]), 
         main = "CDF of 95% Bootstrapped t-stat")
    
  }
}
```

