---
title: "Problem Sheet 1"
author: "Alexander Ober"
date: "2/8/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code is posted at the following site: https://github.com/alexanderober/BUSI525/tree/main/PS1

## Problem 1

Note throughout I test at 5% confidence and as if the alternative is two-sided, since this is what the Wermers survey (pg. 149) suggests to do. I then assume skill is assumed if the t-stat is positive AND significant. In effect, this is a 2.5% one-sided test. Using one-sided 5% tests directly produces similar results. 

### (a)

```{r, echo = FALSE}
## generate variables and simulated data
N <- 1000
T <- 120
mean_mkt <- 0.05/12
sd_mkt <- 0.2/sqrt(12)
sd_res <- 0.1/sqrt(12)
alpha <- rep(0, N)

M <- matrix(rep(rnorm(T, mean = mean_mkt, sd = sd_mkt), N), nrow=T)
M_res <- matrix(rnorm(N*T, mean = 0, sd = sd_res), nrow=T)
M_alpha <- matrix(rep(alpha, T), nrow = T)

## simulate returns in a T by N matrix. Each column is a fund.
M_ret <- M_alpha + M + M_res

## initialize alpha estimates, pvalues, and tstats
alpha_hat = rep(0, N)
pvals = rep(0, N)
tstats = rep(0, N)

## run a regression for each fund 
for (i in 1:N){
  Y <- M_ret[, i]
  X <- M[, i]
  
  mod <- lm(Y~X)
  summ <- summary(mod)
  
  alpha_hat[i] <- summ$coefficients[1, 1]
  tstats[i] <- summ$coefficients[1, 3]
  pvals[i] <- summ$coefficients[1, 4]
  
  ## code to generate one-sided p-value (default output is two-sided)
  # pvals[i] <- ifelse(tstats[i] > 0, 
  #                   summary(mod)$coefficients[1, 4]/2,
  #                   1 - summary(mod)$coefficients[1, 4]/2)
                     
}

```
`r length(pvals[(pvals < 0.05) & (tstats > 0)])` of the 1000 funds have significantly positive alphas. We'd expect about 2.5% = 5/2% of the statistics to be significant and positive, which is what we see.

### (b)
Here are some plots:

```{r, echo = FALSE}
hist(pvals, breaks = 10)
hist(tstats, breaks = 16)
```

### (c)
The distribution looks essentially uniform[0, 1]. This is not too surprising, as under the null the coefficients have a t distribution centered at 0, so p values are uniformly distributed. This is illustrated in one of the footnotes in the Wermers 2011 survey.

## Problem 2

### (a)

```{r, echo = FALSE}
j <- 0
## intialize the true positive, false negative, false positive and true negative 
## vectors for part (b)
tp <- rep(0, 5)
fn <- rep(0, 5)
fp <- rep(0, 5)
tn <- rep(0, 5)
num <-  rep(0, 5)
for (lambda in c(0.1, 0.25, 0.5, 0.75)){
  
  ## generate a proportion of positive alphas
  alpha <- c(rep(0.05/12, lambda*N), rep(0, (1-lambda)*N))
  M_alpha <- matrix(rep(alpha, T), nrow = T, byrow = TRUE)
  
  ## skilled funds correspond to the first lambda*N columns of the matrix
  M_ret <- M_alpha + M + M_res
  
  alpha_hat = rep(0, N)
  pvals = rep(0, N)
  tstats = rep(0, N)
  
  ## run regressions for each fund
  for (i in 1:N){
    Y <- M_ret[, i]
    X <- M[, i]
    
    mod <- lm(Y~X)
    summ <- summary(mod)
    
    alpha_hat[i] <- summ$coefficients[1, 1]
    tstats[i] <- summ$coefficients[1, 3]
    pvals[i] <- summ$coefficients[1, 4]
  
  ## code to generate one-sided p-value (default output is two-sided)
  # pvals[i] <- ifelse(tstats[i] > 0, 
  #                   summary(mod)$coefficients[1, 4]/2,
  #                   1 - summary(mod)$coefficients[1, 4]/2)
    
  }
  
  ## generate some histograms
  title = paste0("Histogram of Alphas: lambda = ", lambda)
  hist(alpha_hat, breaks = 10, main = title)
  title = paste0("Histogram of p-values: lambda = ", lambda)
  hist(pvals, breaks = 20, main = title)
  title = paste0("Histogram of t-stats: lambda = ", lambda)
  hist(tstats, breaks = 16, main = title)
  
  ## compute false positives and negatives
  skilled_pvals <- pvals[1:(lambda*N)]
  skilled_tstats <- tstats[1:(lambda*N)]
  
  unskilled_pvals <- pvals[(lambda*N+1):N]
  unskilled_tstats <- tstats[(lambda*N+1):N]
  
  j <- j+1
  
  tp[j] <- length(skilled_pvals[(skilled_pvals < 0.05) & 
                                  (skilled_tstats > 0)])
  fn[j] <- lambda*N - tp[j]
  fp[j] <- length(unskilled_pvals[(unskilled_pvals < 0.05) & 
                                    (unskilled_tstats > 0)])
  tn[j] <- (1-lambda)*N - fp[j]
  
  ## number of insignificant skilled estimates
  num[j] <-  lambda*N - length(skilled_pvals[(skilled_pvals < 0.05)])
}
```

The distributions of t-stats and p-values visibly shift to the right and left, respectively, as we increase lambda, which is what we'd expect as more of the funds are skilled. This shift in the p-values is particularly pronounced among significant (<5%) p-values. 

### (b)

Below I construct four tables containing proportions (out of N = 1000) of the corresponding samples. 

The number of truly skilled funds with insignificant alpha estimates for $\lambda = 0.1, 0.25, 0.5, 0.75$ are `r num[1]`, `r num[2]`, `r num[3]`, `r num[4]`, respectively. Note, theoretically, these values might differ slightly from the corresponding number in the false negative cell below as the number of insignificant estimates doesn't include significant estimates with negative alphas. In the table, such estimates are identified as unskilled. These numbers account for roughly 60-70% of skilled funds. 

The number of truly unskilled funds identified as skilled based on significantly positive alpha estimates is given in the false positive cell, and generally consists of around 2.5\% of the unskilled funds. Note since the other half of the significant funds are identified as unskilled, this number is 2.5, not 5. 

Lambda = 0.1:
\begin{center}
\begin{tabular}{ c|cc|c } 
 
 Truth\textbackslash Estimate & Positive & Negative &Total \\ 
 \hline
 True & `r tp[1]/N` & `r fn[1]/N` & `r tp[1]/N + fn[1]/N`\\ 
 False & `r fp[1]/N` & `r tn[1]/N` & `r fp[1]/N + tn[1]/N`\\ 
 \hline 
 Total & `r tp[1]/N + fp[1]/N` & `r fn[1]/N + tn[1]/N`&
\end{tabular}
\end{center}

Lambda = 0.25:
\begin{center}
\begin{tabular}{ c|cc|c } 
 
 Truth\textbackslash Estimate & Positive & Negative &Total \\ 
 \hline
 True & `r tp[2]/N` & `r fn[2]/N` & `r tp[2]/N + fn[2]/N`\\ 
 False & `r fp[2]/N` & `r tn[2]/N` & `r fp[2]/N + tn[2]/N`\\ 
 \hline 
 Total & `r tp[2]/N + fp[2]/N` & `r fn[2]/N + tn[2]/N`&
\end{tabular}
\end{center}

Lambda = 0.5:
\begin{center}
\begin{tabular}{ c|cc|c } 
 
 Truth\textbackslash Estimate & Positive & Negative&Total \\ 
 \hline
 True & `r tp[3]/N` & `r fn[3]/N` & `r tp[3]/N + fn[3]/N`\\ 
 False & `r fp[3]/N` & `r tn[3]/N` & `r fp[3]/N + tn[3]/N`\\ 
 \hline 
 Total & `r tp[3]/N + fp[3]/N` & `r fn[3]/N + tn[3]/N`&
\end{tabular}
\end{center}

Lambda = 0.75:
\begin{center}
\begin{tabular}{ c|cc|c } 
 
 Truth\textbackslash Estimate & Positive & Negative&Total \\ 
 \hline
 True & `r tp[4]/N` & `r fn[4]/N` & `r tp[4]/N + fn[4]/N`\\ 
 False & `r fp[4]/N` & `r tn[4]/N` & `r fp[4]/N + tn[4]/N`\\ 
 \hline 
 Total & `r tp[4]/N + fp[4]/N` & `r fn[4]/N + tn[4]/N`&
\end{tabular}
\end{center}
